Proposal for improving content discovery through combining linked data and data mining techniques

### 2015-11-06 proposal/proposal.tex
Tidied up ahead of starting main report.

@@

\title{Proposal for advanced content discovery by a combination of linked data
and data mining techniques}

@@

\section{Introduction}

Media companies produce ever larger numbers of articles, videos, podcasts,
games, etc. -- to which it is commonly referred as ``content''. A successful
content-producing website not only has to develop systems to aid producing and
publishing that content, but there are also calls to engineer effective
mechanisms to aid consumers in finding that content.

@@

\item Many content items are in fairly opaque formats, e.g video content may be
stored as audio-visual binary data with minimal metadata to display on a
containing web page. Video content producers may not be motivated to provide
data attributes that might ultimately be most useful in determining if a user
will enjoy the video.

\item Even with more accessible formats such as text content we need information
retrieval and natural language techniques to provide searchable data structures.
There is not a clear leap from natural language content to more sophisticated
formats suitable for more advanced data mining or machine learning approaches.

@@

@@

In this proposal, I offer the following hypotheses for research:

\item Research and software tools around the concept \emph{Linked Data} can aid
us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
organisation's content whilst also providing a platform for simple enrichment of
that content's metadata.

\item We can establish at least a naïve mapping of an RDF graph of content items
and other entities of interested. With such a mapping, we can explore applying
machine learning -- particularly unsupervised learning -- across an
organisation's whole content corpus in a relatively short time frame.

\item This rapid ease of both acquiring data and providing initial insights with
established algorithms is particularly in line with the principles of the
\emph{Agile} and \emph{Lean} methodologies and we can thus hope to engineer a
prototype-quality web page visualisation of the outcomes from initial data
mining attempts. Such a prototype is arguably a suitable first iteration for a
new content discovery portal, application or website feature an organisation
might wish to develop and release to its visitors.

\section{Outcomes}

The proposed outcomes are outlined here in stages to allow the possibility of
only a subset thereof being achieved in the given time.

>>>>

\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage[autosize]{dot2texi}
\usepackage{xcolor}

@@

\title{Proposal for improving content discovery through combining linked data
and data mining techniques}

@@

\section{Context}

Media companies produce ever larger numbers of articles, videos, podcasts,
games, etc. -- commonly collectively known as ``content''. A successful
content-producing website not only has to develop systems to aid producing and
publishing that content, but there are also demands to engineer effective
mechanisms to aid consumers in finding that content.

@@

\item Many content items are in fairly opaque formats, e.g. video content may be
stored as audio-visual binary data with minimal metadata to display on a
containing web page. Video content producers may not be motivated to provide
data attributes that might ultimately be most useful in determining if a user
will enjoy the video.

@@

\section{Hypotheses}

In this proposal, the following hypotheses are proposed for research:

\item Research and software tools around the concept of \emph{Linked Data} can aid
us in rapidly acquiring a broad view (perhaps at the expense of depth) of an
organisation's content whilst also providing a platform for simple enrichment of
that content's metadata.

\item We can establish at least a na\"ive mapping of an RDF graph representing a
content item to an attribute set suitable for data mining. With such a mapping,
we can explore applying machine learning -- particularly unsupervised learning
-- across an organisation's whole content corpus.

\item Linked Data and Semantic Web \emph{ontologies} and models available can
provide data enrichment beyond attributes and keywords explicitly avaiable
within content data or metadata.

\item Many content-producers currently enrich their web pages with small
amounts of semantic metadata to provide better presentation of that content
as it is shared on social media. This enables simple collection of a full
breadth of content with significantly less effort than direct integration
with content management systems.

@@

It is thus proposed to do the following research activities:

\begin{enumerate}

\item Create a simple software system that will fetch a content item by its
URI, extract semantic data readily available in the page (RDFa, microformats,
social media metadata) and store an RDF graph.

\item Define a simple mapping from these RDF graphs and experiment with
unsupervised learning such as clustering. It should be possible to create a
basic web page that visualises clusters of content as they emerge.

\item Experiment further with the effectiveness of each of the following:
\begin{itemize}
\item Using established entity extraction on textual content to generate new
attributes for a content item's RDF graph.
\item Using linguistic knowledge from taxonomies (e.g. hypernyms) to infer
further attributes that are implicit (i.e. intuitive to humans, but not
explicitly present in the content).
\item Using established ontologies and linked data to enrich a content item
with transitive attributes from related entities (e.g. the birth place of an
actor that appeared on a television programme).
\item Identifying opportunities to apply domain knowledge from experts.
\end {itemize}

\end{enumerate}

A desired outcome is an ecosystem of software services that can receive
notifications of newly-published content, produce highly-enriched RDF graphs,
perform real-time data mining and drive a single web page that allows navigation
of any machine learning insights.

That there is little in the literature around combining linked data and machine
learning indicates that this research could point to the need for deeper
exploration of one or more of the research activities above. It is hoped that
this work will at least provide a good indication of which approaches appear to
give the best opportunties for content metadata enrichment and insights.

It is also hoped that this work provides a good balance of opening up further
research opportunities around data mining of linked data graphs along with some
practical tools and techniques that media organisations could apply immediately
to get a little more effective at providing navigation and discovery of their
content.
